\documentclass[11pt]{article}
  
%%% LATEX COMMON Packages 

\usepackage[bookmarks,colorlinks,breaklinks]{hyperref}  % PDF hyperlinks, with coloured links
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue} % all blue links
\usepackage{amsmath,amsfonts,amsthm,amssymb,xspace,bm,environ,xstring}
\usepackage{dsfont}
\usepackage{fullpage}
\usepackage{mdframed}
\newcommand{\ignore}[1]{}


\title{\bf{Homework 1. Due February 1}}
\author{ CS180: Algorithms and Complexity\\Winter 2017}
\date{}
\begin{document}
\maketitle
\begin{mdframed}
\textsc{Guidelines}:
\begin{itemize}
\item Upload your assignments to Gradescope by 6:59 PM. 
\item Follow the instructions mentioned on the course webpage for uploading to Gradescope very carefully (including starting each problem on a new page and matching the pages with the assignments); this makes it easy and smooth for everyone. As the guidelines are simple enough, bad uploads will not be graded. 
\item You may use results proved in class without proofs as long as you state them clearly.
\item Most importantly, make sure you adhere to the policies for academic honesty set out on the course \href{http://cs180.raghumeka.org}{webpage}. The policies will be enforced strictly. Homework is a stepping stone for exams; keep in mind that reasonable partial credit will be awarded and trying the problems will help you a lot for the exams.
\end{itemize}
\end{mdframed}


\begin{enumerate}
\item Problem 2.8 from \href{https://people.eecs.berkeley.edu/~vazirani/algorithms/chap2.pdf}{[DPV]}. [.75 points]

\item Run the BFS algorithm for the following graph with $s = 1$ and $t = 9$: $G = (V,E)$, where $V = \{1,2,3,4,5,6,7,8,9\}$, and $E = \{\{1,2\}, \{1,3\}, \{2,3\}, \{2,4\},\{2,4\},\{2,5\},\{3,5\},\{3,7\},\{4,5\},\{5,6\}, \{8,9\}\}$. Your work should the step-by-step evolution of the lists $L_0, L_1,...$ and of the array $DISCOVERED$ as we did in class (cf. the transcript uploaded in the calendar). [.75 points]

\item Given a graph $G = (V,E)$ in adjacency list representation, give an algorithm that runs in time $O(|V| \cdot |E|)$ to check if $G$ has a 'triangle', i.e., a triple of distinct vertices $\{u,v,w\}$ such that all three edges between them are present in $G$. [.75 points]

\item Give an algorithm based on BFS that given a graph $G = (V,E)$ (in adjacency list representation) checks whether or not $G$ has a cycle. For full-credit, your algorithm should run in time $O(|V| + |E|)$. Prove that your algorithm works (you can use properties of BFS that we stated in class without further proving them). [.75 points]

\end{enumerate}

\textsc{Additional Problems}. DO NOT turn in answers for the following problems - they are meant for your curiosity and understanding.
\begin{enumerate}
\item Problems 3.4, 3.6, 3.9, 3.11 from textbook [KT].
\item Problem 3.16 from Chapter 3 of \href{https://people.eecs.berkeley.edu/~vazirani/algorithms/chap3.pdf}{[DPV]}.
\end{enumerate}
\end{document}
